---
title: "Project 1: R Component - Titanic Data Exploration"
author: "**Name:** Val Wong - vmw170030"
date:   "**Date:** 3/7/2021"
output:
  html_document:
    df_print: paged
---

## ---Exploration Functions---

### Loading the Titanic data.
```{r}
# setwd("C:\\Users\\Val Wong\\Documents\\RStudio-Workspace\\Proj01")
Titanic <- read.csv("titanic_project.csv") # read csv file
Titanic$pclass <- factor(Titanic$pclass)            # Levels 1,2,3
Titanic$survived <- factor(Titanic$survived)        # Levels 0 1
Titanic$sex <- factor(Titanic$sex)                  # Levels 0 1
```

### Display Column Names and Structure of Titanic Data
Displaying the overall structure of the csv file data frame. 5 variables and their types.
```{r}
names(Titanic) # data exploration function
str(Titanic) # data exploration function
```

### Number of Observations
1046 total rows.
```{r}
print(paste("Total Number of Rows in Titanic: ", NROW(Titanic))) # data exploration function
```

### Preview First and Last Few Rows
Display a couple of rows in the front and end of the Titanic data frame.
```{r}
head(Titanic) # data exploration function
tail(Titanic) # data exploration function
```

### By-Column Analysis
Statistic information of each column in the data frame.
```{r}
summary(Titanic) # data exploration function
```

### Interesting Information
Display total survived = 0 and total survived = 1.
```{r}
print(paste("Total Not Survived (survived = 0): ",sum(Titanic$survived==0))) # data exploration function
print(paste("Total Survived (survived = 1): ",sum(Titanic$survived==1))) # data exploration function
```

## ---Data Visualization---

### Survived
Histogram of Titanic Survived. More values of survived = 0, than survived = 1.
```{r}
plot(factor(Titanic$survived),
     main= "Titanic Data by Survived",
     xlab= "0 = Perished, 1 = Survived")
```

### Sex
More males than females.
```{r}
plot(factor(Titanic$sex),
     main= "Titanic Data by Sex",
     xlab= "0 = Female, 1 = Male")
```

### Passenger Class
More people of passenger class 3.
```{r}
plot(factor(Titanic$pclass),
     main= "Titanic Data by Passenger Class",
     xlab= "Class 1, 2, or 3")
```

### Age
Age visual statistic information. Age 18 to 25 are higher in distribution.
```{r}
hist(Titanic$age,
     main= "Titanic Data by Age",
     xlab= "age (in yrs)")
```

### Passenger Class vs. Survived
PClass vs Survived. More of passenger class 2 and passenger class 3 perished than survived. Passenger class 1 survived way more compared to other classes.
```{r}
plot(as.factor(Titanic$pclass), as.factor(Titanic$survived),
     main= "Passenger Class vs. Survived",
     xlab= "Passenger Class",
     ylab= "0 = Perished, 1 = Survived")

```

### Passenger Class vs. Sex
Most female survived and most male perished.
```{r}
plot(as.factor(Titanic$sex), as.factor(Titanic$survived),
     main= "Sex vs. Survived",
     xlab= "0 = Female, 1 = Male",
     ylab= "0 = Perished, 1 = Survived")
```
### GLM Residual Plot
Plot of the residuals vs. regression line.
```{r}
i <- 1:900
train <- Titanic[i, ] # 900 in train
test <- Titanic[-i, ] # 146 in test
glm1 <- glm(survived ~ pclass, data=train, family="binomial")
plot(glm1)
```

### Overall
Titanic data frame (as a whole) plot
```{r}
plot(Titanic)
```

